Choose a License

### 2025-02-08 archive.qmd
Recommend against OSF for archival
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

The project folder is now ready to be deposited for long-term archival. When considering an appropriate repository, there are a few questions to ponder [excerpt from @Jackson2018, with footnotes removed, licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)]:

> __Does the digital repository give you a unique persistent digital identifier for your deposit?__
> 
> Persistent digital identifiers are identifiers that are provided under the assumption that they, and the digital objects they refer to, will remain available long term. Examples of unique persistent digital identifiers include DOIs and ARKs. The use of persistent digital identifiers for research outputs, including papers, data and software, and the use of these for citation purposes, is becoming prevalent in research.
> 
> __Can the digital repository accommodate the size of your deposit?__
> 
> Digital repositories may have constraints on the size of deposits they accept. For example, Zenodo has a limit of 50 GB and the University of Edinburgh DataShare has a limit of 20 GB for deposits done via a browser and 100 GB otherwise.
> 
> __Has your funder or publisher recommended or mandated a digital repository to use?__
> 
> If both a funder and publisher mandate two different digital repositories you could deposit within them both, or discuss with each to see if they will agree for you to use one only, even if it is not the one they mandate.
> 
> __Are the policies of the digital repository acceptable to you?__
> 
> These can include terms and conditions of use; availability of the service (e.g. does the service regularly go down for updates or does it run 24/7); how resilient the service is to breakdowns; where deposits are stored; how they are stored; what are the plans for stored deposits if the repository goes out of business; how secure the storage is; how regularly it is backed up; how long backups are kept for; and what disaster recovery plans are in place. Knowing the quality of service can help you judge whether your deposit will be available to others, and yourself, even in the face of disaster. It also allows you, and your institution, to develop contingency plans should such drastic circumstances arise.
> 
> __Is the longevity of the digital repository acceptable to you?__
> 
> You want to know that the digital repository will host your deposit for as long as you need it too, not just 6 months or so. The policies or terms and conditions of the digital repository should state this. For example, Zenodo states that their lifetime is currently 20 years and figshare guarantees 10 years of persistent availability.
> 
> __Is the digital repository free or do you have to pay a fee?__
> __If there is a fee, is this a one-off payment and can you afford it?__
> 
> If a digital repository expects a one-off payment, to subsidise its hosting, you may be able to pay for this out of your project budget or your institution may pay this for you. It is recommended that you avoid any digital repository that expects a regular fee for the duration for which they hold the deposit, since it would most likely be unsustainable for your institution to commit to ever-increasing fees over a long term. There are many free digital repositories available.
> 
> __Is the digital repository accredited or certified?__
> 
> The digital repository may be accredited or certified by an independent agency that reviews digital repositories according to various criteria related to its policies and processes. The presence of an accreditation or certification may provide you with an additional degree of confidence that the digital repository meets your requirements. For example, the Data Seal of Approval can be awarded to digital repositories who satisfy a set of quality guidelines relating to how they manage, preserve and make available the deposits they hold. The presence of a Data Seal of Approval could give you additional confidence that a digital repository is suitable for holding your deposit.

Additional guidance is provided by @Daquin2023, @Garijo2022, and @ScienceEurope2021. You can use the registry of research data repositories [re3data.org](https://www.re3data.org/) to find one that matches your criteria. We think that the [Software Heritage code archive](https://www.softwareheritage.org/) and [Zenodo](https://zenodo.org/) are good choices, but don't forget to also check your local institutions for recommendations.

@@

Note that during this tutorial, we will be using Zenodo's sandbox instance located at [`sandbox.zenodo.org`](https://sandbox.zenodo.org/) in order to learn how to use it while not wasting Zenodo's resources. For real projects, use the production instance at [`zenodo.org`](https://zenodo.org/), which requires a separate account.
:::

@@

Now you can follow the guide "[Create new upload](https://help.zenodo.org/docs/deposit/create-new-upload/)", taking into account the following notes specific to each step:

>>>>

The project folder is now ready to be deposited for long-term archival.
When considering an appropriate repository,
there are a few questions to ponder
[excerpt from @Jackson2018, with footnotes removed, licensed under [CC\ BY\ 4.0][cc-by]]:

[cc-by]: https://creativecommons.org/licenses/by/4.0/

> __Does the digital repository give you a unique persistent digital identifier
> for your deposit?__
> 
> Persistent digital identifiers are identifiers that are provided under the
> assumption that they, and the digital objects they refer to, will remain
> available long term. Examples of unique persistent digital identifiers include
> DOIs and ARKs. The use of persistent digital identifiers for research outputs,
> including papers, data and software, and the use of these for citation
> purposes, is becoming prevalent in research.
> 
> __Can the digital repository accommodate the size of your deposit?__
> 
> Digital repositories may have constraints on the size of deposits they accept.
> For example, Zenodo has a limit of 50 GB and the University of Edinburgh
> DataShare has a limit of 20 GB for deposits done via a browser and 100 GB
> otherwise.
> 
> __Has your funder or publisher recommended or mandated a digital repository
> to use?__
> 
> If both a funder and publisher mandate two different digital repositories you
> could deposit within them both, or discuss with each to see if they will agree
> for you to use one only, even if it is not the one they mandate.
> 
> __Are the policies of the digital repository acceptable to you?__
> 
> These can include terms and conditions of use; availability of the service
> (e.g. does the service regularly go down for updates or does it run 24/7);
> how resilient the service is to breakdowns; where deposits are stored;
> how they are stored; what are the plans for stored deposits if the repository
> goes out of business; how secure the storage is; how regularly it is backed up;
> how long backups are kept for; and what disaster recovery plans are in place.
> Knowing the quality of service can help you judge whether your deposit will be
> available to others, and yourself, even in the face of disaster. It also
> allows you, and your institution, to develop contingency plans should such
> drastic circumstances arise.
> 
> __Is the longevity of the digital repository acceptable to you?__
> 
> You want to know that the digital repository will host your deposit for as
> long as you need it too, not just 6 months or so. The policies or terms and
> conditions of the digital repository should state this. For example, Zenodo
> states that their lifetime is currently 20 years and figshare guarantees
> 10 years of persistent availability.
> 
> __Is the digital repository free or do you have to pay a fee?__
> __If there is a fee, is this a one-off payment and can you afford it?__
> 
> If a digital repository expects a one-off payment, to subsidise its hosting,
> you may be able to pay for this out of your project budget or your institution
> may pay this for you. It is recommended that you avoid any digital repository
> that expects a regular fee for the duration for which they hold the deposit,
> since it would most likely be unsustainable for your institution to commit to
> ever-increasing fees over a long term. There are many free digital
> repositories available.
> 
> __Is the digital repository accredited or certified?__
> 
> The digital repository may be accredited or certified by an independent agency
> that reviews digital repositories according to various criteria related to its
> policies and processes. The presence of an accreditation or certification may
> provide you with an additional degree of confidence that the digital
> repository meets your requirements. For example, the Data Seal of Approval can
> be awarded to digital repositories who satisfy a set of quality guidelines
> relating to how they manage, preserve and make available the deposits they hold.
> The presence of a Data Seal of Approval could give you additional confidence
> that a digital repository is suitable for holding your deposit.

Additional guidance is provided by @Daquin2023, @Garijo2022, and @ScienceEurope2021.
You can use the registry of research data repositories [re3data.org][re3data]
to find one that matches your criteria.
We think that the [Software Heritage code archive][sw-heritage]
and [Zenodo][zenodo] are good choices,
but don't forget to also check your local institutions for recommendations.
Note that the Open Science Framework (OSF) is not recommended
because it allows the deletion of files and projects after their publication.

[re2data]: https://www.re3data.org/
[sw-heritage]: https://www.softwareheritage.org/
[zenodo]: https://zenodo.org/

@@

Note that during this tutorial, we will be using Zenodo's sandbox instance
located at [`sandbox.zenodo.org`][zenodo-sandbox] in order to learn
how to use it while not wasting Zenodo's resources.
For real projects, use the production instance at [`zenodo.org`][zenodo],
which requires a separate account.
:::

[zenodo-sandbox]: https://sandbox.zenodo.org/

@@

Now you can follow the guide "[Create new upload][zenodo-upload]",
taking into account the following notes specific to each step:

[zenodo-upload]: https://help.zenodo.org/docs/deposit/create-new-upload/

### 2025-02-08 choose_license.qmd
Reformat markdown
To ease change tracking in the future.
Too bad it wasn't done from the
beginning.

We recommend to choose free/open licenses suitable to the type of work (e.g., code, data, image etc.). [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) and [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) are good choices to apply simultaneously to all own works. Take your time to read the summaries behind the links and/or the actual license text to understand the legal effect. You can adapt the following wording to your use case:

> Except where noted otherwise, all files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache License\ 2.0.
:::

@@

::: {.column-margin}
If you would like to choose a license not listed here, it should be appropriate for the type of work in question and be compatible with the dominant copyleft license in the respective community (see also @nte-silos). For software that's almost universally the (A)GPLv3 and for writing, image, audio, and video that's mostly the [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/). For databases, no dominant copyleft license has emerged yet, so any of [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/), [CDLA\ Sharing\ 1.0](https://cdla.dev/sharing-1-0/), and [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) are acceptable.
:::

@@

First, if you adapt (i.e., modify, remix, build on) a work by others you need to determine if it is provided to you under a free/open license. If yes, we recommend you to make your contribution available under the same license.^[Copyleft licenses even require you to choose the same or a compatible license.] For example, if you adapt code published in another paper, choose the same license for your modifications. The same applies if there are strong community norms to use a particular free/open license^[Of course, this is only a heuristic and there might be good reasons to deviate from community norms.] or if a particular license choice is mandated by your funder. Importantly, as discussed before, you are generally not allowed to adapt a work _not_ published under a free/open license.

@@

One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights [see @CC2024Software]] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to display the full text of the license [see @Wikipedia2024FDL]]

One solution is to make such a work simultaneously available under two (or more) licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video]. This is called multi-licensing and makes it easier to reuse both the rendered document as well as the code. For example, one could write:

> The Quarto documents in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache License\ 2.0.
:::

We already mentioned that a database may be licensed differently than its content. However, if the content was created by you, we recommend you to choose the same license for both content and database. Factual data (like measurements or metadata) should be made available with [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) [@Villa2016Copyleft; @Villa2016CC0] -- otherwise, consult @nte-other-work-types for some caveats.

@@

- __fonts:__ Copyleft licenses applied to fonts can be a special case: If a font is put under the license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), any documents containing texts using that font will probably be derivative works and have to be put under the same license if shared. The [SIL Open Font License\ 1.1](https://openfontlicense.org/) is the _de facto_ standard license for fonts, requiring that derived fonts, if published, have to be put under the same license. It doesn't require attribution for usage (e.g., when embedded in a document), but forbids that the font is sold by itself. Selling the font is allowed, however, if it is bundled with other software.
- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package, is licensed under a copyleft license such as the (A)GPLv3, every work that is a derivative of it has to be put under the same license if shared. If a document contained source code covered by the (A)GPLv3, it depends on the individual case whether the same license needs to be applied to the document if shared.
- __database content:__ If the work produced from a database (the "output") is a derivative of the content in the database, the output is subject to the restrictions laid out in the license. For example, if geospatial data were to be licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/), all maps produced from the data would likely need to fulfill this license's obligation for unrestricted access (its Anti-DRM provision) if shared [see @Poole2017]. Similarly, following an example from @Matt2009, if one were to choose the copyleft license [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) for this purpose, any map that is a derivative of the data would also need to be licensed under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/) (or a compatible license) if shared. If the intention is to only have [derivative databases](https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline) under the same license, one might want to choose the [ODbL\ 1.0](https://opendatacommons.org/licenses/odbl/summary/) for the database, as it was specifically designed not to apply to works produced from the data in the database. Otherwise [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) is an excellent choice for data.
- __works in the public domain__: If a work is already in the public domain, it should be marked using the [PDM 1.0](https://creativecommons.org/publicdomain/mark/1.0/), rather than applying a waiver such as the [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) (or another license).
:::

@@

In most cases, the output of software, like images or tables, does not depend on the software's license.^[The exception is when the software is part of the output [see @FSF2024].] Therefore, if you use an R package under a copyleft license to create a figure, you are likely the copyright owner. However, if the output is based on data, it can be considered a derivative work of the data and the license of the data also applies. For example, maps may be considered as a derivative of the geographic data they are based on.

It is disputed whether software that uses an R package under AGPLv3 or GPLv3^[a variant of the AGPLv3 that does not cover software running as a service] can only be published under a GPL-compatible license -- or even has to be published under the same license. Posit, the company behind RStudio, does not believe that to be the case [see also @Wickham2023PackagesLicense].

@@

From the Creative Commons licenses, only [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) does not require providing attribution, and all software licenses in @fig-flowchart-non-data require providing attribution. As an author, you are free to waive license conditions such as attribution, however.

@@

!["Open data-capable license interoperability" by Robbie Morrison [-@Morrison2024] licensed under [CC\ BY\ 4.0](https://creativecommons.org/licenses/by/4.0/)](images/Morrison2024.jpg){.lightbox}
:::

@@

- Applying the license is irrevocable. You may stop distributing your work, but everybody who already received it may continue to share it.
- As an author applying the license, you are not bound by the license terms. You cannot, however, give others an exclusive license to your content anymore (as some journals needlessly demand, see @wrn-exclusive-rights).
- By applying the license you share with others your right to use your work for commercial purposes, including selling it.
- If you apply the license to an image, the license [may also apply](https://commons.wikimedia.org/wiki/Commons:Same_work) to higher resolution version of your image.
- If you apply a copyleft license to an image, a document in which the image is used may not constitute a derivative work and, as such, may not be affected by the copyleft condition.
- If you are a member of a collecting society^[in Germany: _Verwertungsgesellschaft_] or if the work was created as part of your job, you may not be able to apply the license. Check with your collecting society or your employer in that case.

@@

- You have already learned about the file `LICENSE.txt`. Either it contains the full text of one license that applies to the whole project folder -- this is a practice propagated by GitHub, which provides instructions for and comparisons of many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/). Alternatively, it details which license applies to what and the invidual licenses are stored in the same or in different files.
- A summary of the licensing situation is often given in a `README` file, whose creation will be discussed [later](make_readme.qmd).
- If only parts of a file are covered by a certain license, then it might make sense to indicate this directly in the file.
- The license of embedded elements (e.g., images) is either given directly beneath to where they are embedded, at the end of the respective page, or collectively on an separate page.
- Individual programming languages also have their own way of stating which license a package is distributed under. For R packages, this is usually set by the field `License` in the file `DESCRIPTION` [@Wickham2023PackagesLicense].
- Software projects often indicate the license in every file, following the [REUSE](https://reuse.software/) specification to make the choice of license machine-readable (explained in @nte-reuse).
- The following two metadata standards for increasing citability also allow to indicate the license:
  - [CodeMeta](https://codemeta.github.io/): All information are stored in a file called `codemeta.json`, which can be created [on their website](https://codemeta.github.io/codemeta-generator/). Alternatively, one can use the R package [`codemetar`](https://docs.ropensci.org/codemetar/) to create this file.
  - [Citation File Format](https://citation-file-format.github.io/) (CFF): Here, everything is stored in a file called `CITATION.cff`. It is used by GitHub to directly show a recommended citation for a repository. The file can be created [on their website](https://citation-file-format.github.io/cff-initializer-javascript/). Alternatively, the R package [`cffr`](https://cran.r-project.org/package=cffr) may be used for that purpose.

@@

Note that `Manuscript.qmd` was provided to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/), which is what you indicate with the previous comment. As you edited the file throughout this tutorial, you may also add yourself.^[For licenses that require that modifications are indicated, this is an easy way to comply with them. Also, you do not need to provide your real name.]

Sometimes, there are file types which do not allow for adding the license information inside them, such as PDF and CSV files. For these, a corresponding `.license` file can be created. Try the following command which indicates that the data were published under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/):

@@

The manuscript and the bibliography file you downloaded for this tutorial are made available to you under [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) as is all code in this tutorial (e.g., for creating the data dictionary). Therefore, you can apply any license to your modified version. Please apply the license of your choice now, using any of the methods mentioned above (except the `README`, which we will cover later).

@@

In the following example, the edited manuscript files have been multi-licensed under [AGPLv3](https://www.gnu.org/licenses/agpl-3.0.html) (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0](https://creativecommons.org/licenses/by-sa/4.0/), while the bibliography, the data dictionary and its code are made available under CC0\ 1.0.

>>>>

We recommend to choose free/open licenses suitable to the type of work
(e.g., code, data, image etc.).
[CC0\ 1.0][cc0] and [Apache\ 2.0][apache2] are good choices
to apply simultaneously to all own works.
Take your time to read the summaries behind the links
and/or the actual license text to understand the legal effect.
You can adapt the following wording to your use case:

[cc0]: https://creativecommons.org/publicdomain/zero/1.0/
[apache2]: https://choosealicense.com/licenses/apache-2.0/

> Except where noted otherwise, all files in this project
> are made available under CC0\ 1.0
> or (at your option) under the terms of the Apache License\ 2.0.
:::

@@

::: {.column-margin}
If you would like to choose a license not listed here,
it should be appropriate for the type of work in question
and be compatible with the dominant copyleft license in the respective community
(see also @nte-silos).
For software that's almost universally the (A)GPLv3
and for writing, image, audio, and video that's mostly the [CC\ BY-SA\ 4.0][cc-by-sa].
For databases, no dominant copyleft license has emerged yet,
so any of [ODbL\ 1.0][odbl], [CDLA\ Sharing\ 1.0][cdla-sharing],
and [CC\ BY-SA\ 4.0][cc-by-sa] are acceptable.
:::

[cc-by-sa]: https://creativecommons.org/licenses/by-sa/4.0/
[odbl]: https://opendatacommons.org/licenses/odbl/summary/
[cdla-sharing]: https://cdla.dev/sharing-1-0/

@@

First, if you adapt (i.e., modify, remix, build on) a work by others
you need to determine if it is provided to you under a free/open license.
If yes, we recommend you to make your contribution available
under the same license.[^required-with-copyleft]
For example, if you adapt code published in another paper,
choose the same license for your modifications.
The same applies if there are strong community norms
to use a particular free/open license[^heuristic]
or if a particular license choice is mandated by your funder.
Importantly, as discussed before, you are generally not allowed
to adapt a work _not_ published under a free/open license.

[^required-with-copyleft]: Copyleft licenses even require you to choose the same
or a compatible license.

[^heuristic]: Of course, this is only a heuristic
and there might be good reasons to deviate from community norms.

@@

One may wonder which license to apply in this case,
because Creative Commons licenses are not recommended for source code[^disclaim-patents]
and applying software licenses to PDFs or images
can lead to confusion or nuisance.[^require-full-text]

[^disclaim-patents]: because (among other reasons) they explicitly disclaim
any conveyance of patent rights [see @CC2024Software]

[^require-full-text]: because they often require
to display the full text of the license [see @Wikipedia2024FDL]

One solution is to make such a work simultaneously available
under two (or more) licenses, at the choice of the recipient:
Either under a specified software license,
or under a Creative Commons license[^media].
This is called multi-licensing and makes it easier
to reuse both the rendered document as well as the code.
For example, one could write:

[^media]: that is, a license for writing, image, audio, and video

> The Quarto documents in this project are made available under CC0\ 1.0
> or (at your option) under the terms of the Apache License\ 2.0.
:::

We already mentioned that a database may be licensed differently than its content.
However, if the content was created by you,
we recommend you to choose the same license for both content and database.
Factual data (like measurements or metadata) should be made available
with [CC0\ 1.0][cc0] [@Villa2016Copyleft; @Villa2016CC0]
-- otherwise, consult @nte-other-work-types for some caveats.

@@

- __fonts:__ Copyleft licenses applied to fonts can be a special case:
  If a font is put under the license [CC\ BY-SA\ 4.0][cc-by-sa],
  any documents containing texts using that font will probably be derivative works
  and have to be put under the same license if shared.
  The [SIL Open Font License\ 1.1][ofl] is the _de facto_ standard license for fonts,
  requiring that derived fonts, if published, have to be put under the same license.
  It doesn't require attribution for usage (e.g., when embedded in a document),
  but forbids that the font is sold by itself.
  Selling the font is allowed, however, if it is bundled with other software.
- __templates and TeX packages:__ If a template or, for this purpose, a $\TeX$ package,
  is licensed under a copyleft license such as the (A)GPLv3,
  every work that is a derivative of it has to be put under the same license if shared.
  If a document contained source code covered by the (A)GPLv3,
  it depends on the individual case
  whether the same license needs to be applied to the document if shared.
- __database content:__ If the work produced from a database (the "output")
  is a derivative of the content in the database,
  the output is subject to the restrictions laid out in the license.
  For example, if geospatial data were to be licensed under [CC\ BY\ 4.0][cc-by],
  all maps produced from the data would likely need to fulfill
  this license's obligation for unrestricted access (its Anti-DRM provision)
  if shared [see @Poole2017].
  Similarly, following an example from @Matt2009,
  if one were to choose the copyleft license [CC\ BY-SA\ 4.0][cc-by-sa]
  for this purpose, any map that is a derivative of the data
  would also need to be licensed under [CC\ BY-SA\ 4.0][cc-by-sa]
  (or a compatible license) if shared.
  If the intention is to only have [derivative databases][osm-produced-work]
  under the same license, one might want to choose the [ODbL\ 1.0][odbl] for the database,
  as it was specifically designed not to apply to works produced from the data in the database.
  Otherwise [CC0\ 1.0][cc0] is an excellent choice for data.
- __works in the public domain__: If a work is already in the public domain,
  it should be marked using the [PDM 1.0][pdm],
  rather than applying a waiver such as the [CC0\ 1.0][cc0] (or another license).
:::
https://creativecommons.org/licenses/by/4.0/
[ofl]: https://openfontlicense.org/
[cc-by]: https://creativecommons.org/licenses/by/4.0/
[pdm]: https://creativecommons.org/publicdomain/mark/1.0/
[osm-produced-work]: https://osmfoundation.org/wiki/Licence/Community_Guidelines/Produced_Work_-_Guideline

@@

In most cases, the output of software, like images or tables,
does not depend on the software's license.[^output-exception]
Therefore, if you use an R package under a copyleft license to create a figure,
you are likely the copyright owner.
However, if the output is based on data,
it can be considered a derivative work of the data and the license of the data also applies.
For example, maps may be considered as a derivative of the geographic data they are based on.

[^output-exception]: The exception is when the software is part of the output [see @FSF2024].

It is disputed whether software that uses an R package
under AGPLv3 or GPLv3[^gplv3-explanation]
can only be published under a GPL-compatible license
-- or even has to be published under the same license.
Posit, the company behind RStudio,
does not believe that to be the case [see also @Wickham2023PackagesLicense].

[^gplv3-explanation]: a variant of the AGPLv3 that does not cover software running as a service

@@

From the Creative Commons licenses,
only [CC0\ 1.0][cc0] does not require providing attribution,
and all software licenses in @fig-flowchart-non-data require providing attribution.
As an author, you are free to waive license conditions such as attribution, however.

@@

!["Open data-capable license interoperability" by Robbie Morrison [-@Morrison2024]
licensed under [CC\ BY\ 4.0][cc-by]](images/Morrison2024.jpg){.lightbox}
:::

@@

- Applying the license is irrevocable.
  You may stop distributing your work,
  but everybody who already received it may continue to share it.
- As an author applying the license, you are not bound by the license terms.
  You cannot, however, give others an exclusive license to your content anymore
  (as some journals needlessly demand, see @wrn-exclusive-rights).
- By applying the license you share with others your right
  to use your work for commercial purposes, including selling it.
- If you apply the license to an image,
  the license [may also apply][commons-same-work] to a higher resolution version of your image.
- If you apply a copyleft license to an image, a document
  in which the image is used may not constitute a derivative work
  and, as such, may not be affected by the copyleft condition.
- If you are a member of a collecting society^[in Germany: _Verwertungsgesellschaft_]
  or if the work was created as part of your job,
  you may not be able to apply the license.
  Check with your collecting society or your employer in that case.

[commons-same-work]: https://commons.wikimedia.org/wiki/Commons:Same_work

@@

- You have already learned about the file `LICENSE.txt`.
  Either it contains the full text of one license that applies to the whole project folder
  -- this is a practice propagated by GitHub,
  which provides instructions for and comparisons of many licenses
  via [ChooseALicense.com](https://choosealicense.com/licenses/).
  Alternatively, it details which license applies to what
  and the individual licenses are stored in the same or in different files.

- A summary of the licensing situation is often given in a `README` file,
  whose creation will be discussed [later](make_readme.qmd).

- If only parts of a file are covered by a certain license,
  then it might make sense to indicate this directly in the file.

- The license of embedded elements (e.g., images) is either given
  directly beneath to where they are embedded,
  at the end of the respective page,
  or collectively on an separate page.

- Individual programming languages also have their own way
  of stating which license a package is distributed under.
  For R packages, this is usually set by the field `License`
  in the file `DESCRIPTION` [@Wickham2023PackagesLicense].

- Software projects often indicate the license in every file,
  following the [REUSE](https://reuse.software/) specification
  to make the choice of license machine-readable (explained in @nte-reuse).

- The following two metadata standards for increasing citability
  also allow to indicate the license:

  - [CodeMeta](https://codemeta.github.io/):
    All information are stored in a file called `codemeta.json`,
    which can be created [on their website][codemeta-gen].
    Alternatively, one can use the R package [`codemetar`][codemetar]
    to create this file.

  - [Citation File Format](https://citation-file-format.github.io/) (CFF):
    Here, everything is stored in a file called `CITATION.cff`.
    It is used by GitHub to directly show a recommended citation for a repository.
    The file can be created [on their website][cff-gen].
    Alternatively, the R package [`cffr`][cffr] may be used for that purpose.

[codemeta-gen]: https://codemeta.github.io/codemeta-generator/
[codemetar]: https://docs.ropensci.org/codemetar/
[cff-gen]: https://citation-file-format.github.io/cff-initializer-javascript/
[cffr]: https://cran.r-project.org/package=cffr

@@

Note that `Manuscript.qmd` was provided to you under [CC0\ 1.0][cc0],
which is what you indicate with the previous comment.
As you edited the file throughout this tutorial,
you may also add yourself.^[For licenses that require that modifications are indicated,
this is an easy way to comply with them.
Also, you do not need to provide your real name.]

Sometimes, there are file types which do not allow
for adding the license information inside them, such as PDF and CSV files.
For these, a corresponding `.license` file can be created.
Try the following command which indicates that the data were published
under [CC0\ 1.0][cc0]:

@@

The manuscript and the bibliography file you downloaded for this tutorial
are made available to you under [CC0\ 1.0][cc0]
as is all code in this tutorial (e.g., for creating the data dictionary).
Therefore, you can apply any license to your modified version.
Please apply the license of your choice now,
using any of the methods mentioned above (except the `README`, which we will cover later).

@@

In the following example, the edited manuscript files have been multi-licensed
under [AGPLv3][agpl] (or later) or (at option of the reuser) under [CC\ BY-SA\ 4.0][cc-by-sa],
while the bibliography, the data dictionary and its code are made available under CC0\ 1.0.

[agpl]: https://www.gnu.org/licenses/agpl-3.0.html
